### 2023-03-17 chapter0/chapter0.tex
Added paragraph on SLiMs

The structural diversity of IDRs is directly related to their functional plasticity.

>>>>

The structural diversity of IDRs is directly related to their functional plasticity, and as IDRs are not confined to one conformation, they can interact with and bind to many possible partners. Often these partners are other proteins, but they can also be other macromolecules like DNA or even small molecules and ions. As a result of this adaptability, IDRs are enriched in proteins involved in cell signaling and regulation. Because cells are highly compartmentalized, they have a dizzying array of mechanisms to relay messages.\footnote{The most fundamental compartment is the cell, which roughly separates inside from outside and life from non-life. However, the cells of more complex organisms called eukaryotes contain additional subcompartments called organelles. Compartmentalization is essential in living systems because it confines and concentrates biochemical reactions that would be harmful if they occurred at the wrong place or time. However, it also introduces many complications because information in the form of physical or chemical signals cannot travel freely.} Many are mediated by the interaction of one protein with another, which in turn creates a change in the state of the system that propagates the signal. Often these state changes are chemical alterations made by one protein to another, which are termed ``post-translational modifications'' (PTMs).\footnote{A protein's amino acid sequence is encoded in a cell's DNA, and the process of reading that information to create a protein (from an intermediate molecule called RNA) is translation. Therefore, any modifications to a protein after its initial synthesis are post-translational.} One of the most common modifications is phosphorylation where a phosphate group is attached to a specific amino acid in the protein. Phosphate groups contain three negative charges, so their addition can dramatically affect a protein's structural energetics and induce a conformational change. Thus, phosphorylation often plays a key role in toggling proteins between inactive and active states. In general, though, PTMs encompass a variety of chemical modifications with similarly diverse impacts on a protein's behavior. Because IDRs are generally exposed to their environment, they are frequent targets of PTMs. However, PTMs do not occur haphazardly in proteins but are instead targeted to binding sites created by sequential patterns of residues called short linear motifs (SLiMs). SLiMs are a general mechanism for mediating interactions with proteins, so while many SLiMs are targets of PTMs, others simply recruit binding partners. In contrast to the highly structured interfaces that characterize folded domains, however, SLiMs are usually no more than ten residues. Thus, their interactions are relatively weak and therefore highly transient. As their flexibility makes SLiMs easily accessible to binding partners, they are highly enriched in IDRs.
